Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes

### 2019-12-24 ms/manuscript.Rmd
Revise discussion and acknowledgements

We did not find that gemmae production in gametophytes was linked with epiphytism. 
Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (\fig \@ref(fig:tree)).
Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
In particular, gemmae production should be useful in an epiphytic context because they may allow gametophytes to increase their boundary layer and hedge against population loss during extreme drying [@Farrar2008].
Gemmae can also function sexually by producing antheridia in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
Interestingly, when we classified 'widespread' species as those with gametophytes occurring above or below the minimum or maximum elevational range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with widespread growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].

@@

Our results demonstrate that the canopy is far more dynamic with regard to fern epiphytism than has been previously appreciated. 
On the one hand, the canopy appears to have served as a strong barrier to entry, sorting species by size. 
Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (\fig \@ref(fig:comm-div)).
This makes sense given that most epiphytes have radiated within five major clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.

@@

We thank members of the Davis Lab, N. Michele Holbrook, and Jonathan Losos for helpful discussion and comments on drafts. Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits. Ravahere Taputuarai helped collect environmental data. Staff at the University of California, Berkeley Richard B. Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics. Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and CCD), Setup Funds from Harvard University to CCD, the American Society of Plant Taxonomists (Research Grant for Graduate Students to JHN), the Garden Club of America (Award in Tropical Botany to JHN), the Harvard University Herbaria (Fernald Fieldwork Fellowship to JHN), the Society of Systematic Biologists (Graduate Student Research Award to JHN), and the Systematics Association (Systematics Research Fund to JHN).

>>>>

We did not find that gemmae production in gametophytes was linked with epiphytism. 
Rather, occurrence of gemmae strongly reflects phylogeny, appearing concentrated in filmy ferns (Hymenophyllaceae) and vittarioid ferns (\fig \@ref(fig:tree)).
Gemmae are asexual propagules occurring primarily in non-cordate gametophytes that allow gametophytes to persist over multiple growing seasons and attain large population sizes, and have been hypothesized to be associated with harsh environmental conditions [@Farrar2008].
In particular, gemmae production should be useful in an epiphytic context because they may allow gametophytes to increase their boundary layer and hedge against population loss during extreme drying [@Farrar2008].
Gemmae can also function sexually by producing antheridia in the presence of antheridiogen secreted by other gametophytes [@Emigh1977], which may confer a selective advantage in the epiphytic environment by enabling outcrossing [@Dassler2001].
As with our results summarized above, it is possible that the diversity of gemmae and non-gemmae producing epiphytic species reflects undetected niche variation in the canopy.
Interestingly, when we classified gametophytes as ‘independent’ if they occurred beyond the range of conspecific sporophytes using the data of @Nitta2017, we found that gemmae production is correlated with independent growth (log likelihood dependent model = `r correlated_evo_test_gemmae$dependent.logL %>% round(1) %>% as.numeric`, log likelihood independent model = `r correlated_evo_test_gemmae$independent.logL %>% round(1) %>% as.numeric`, \pval = `r correlated_evo_test_gemmae$P %>% round(3) %>% as.numeric`, Pagel's test of correlated evolution), echoing a pattern observed in Japan of non-cordate gametophytes frequently occurring independently of their conspecific sporophytes [@Ebihara2013].

@@

Our results demonstrate that the canopy is far more dynamic with regard to fern epiphytism than has been previously appreciated. 
On the one hand, the canopy appears to have served as a potent barrier to entry, strongly selecting for small size. 
Furthermore, the results of our phylogenetic community structure analyses partially support the hypothesis of decreased phylogenetic diversity in epiphytic communities due to filtering: we observed that epiphytic fern communities were more phylogenetically clustered for MNTD~phy~, but not MPD~phy~ (\fig \@ref(fig:comm-div)).
This makes sense given that most epiphytes have radiated within a limited number of clades [@Schuettpelz2009]; each epiphytic community we sampled tends to include closely related species from two or more epiphytic clades, but exhibits a similar overall degree of phylogenetic diversity as terrestrial communities.

@@

We thank members of the Davis Lab, N. Michele Holbrook, and Jonathan Losos for helpful discussion and comments on drafts. Three anonymous reviewers provided comments that improved the manuscript. Suzanne Vinette, Saad Amer, Tristan Wang, and Tohei Theophilus helped with community surveys and measurement of traits. Ravahere Taputuarai helped collect environmental data. Staff at the University of California, Berkeley Richard B. Gump South Pacific Research Station, Moorea, French Polynesia, in particular Valentine Brotherson, Hinano Teavai-Murphy, Frank Murphy, and Neil Davies, provided support with obtaining research and collection permits, laboratory space, and logistics. Funding provided in part by National Science Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to J.H.N. and C.C.D.), Setup Funds from Harvard University to C.C.D., the American Society of Plant Taxonomists (Research Grant for Graduate Students to J.H.N.), the Garden Club of America (Award in Tropical Botany to J.H.N.), the Harvard University Herbaria (Fernald Fieldwork Fellowship to J.H.N.), the Society of Systematic Biologists (Graduate Student Research Award to J.H.N.), and the Systematics Association (Systematics Research Fund to J.H.N.).
